---
title: "kNN Mutual Information Estimators"
author: "Isaac J. Michaud"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kNN Mutual Information Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of this work is to introduce basic kNN estimators of mutual information and give a few examples of how to utilize. This document is the draft of an R package vignette.Mutual information is important measures dependence in ways that are not linear like correlations. 

## Mutual Information

Mutual information is an information theoretic quantity that measures the dependence between two random variables by quantifying the amount of information (in terms of Shannon entropy) knowledge of one variable provides about the other. The mutual information between two random variables $X$ and $Y$ is defined by

\begin{equation}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
  (\#eq:binom)
\end{equation}

$$ I(X;Y) =\int \int \log \left \{ \frac{f_{xy}(x,y)}{f_{x}(x)f_{y}(y)} \right \} f_{xy}(x,y) dy dx, $$ 

where $f_x$ and $f_y$ are the marginal densities of $X$ and $Y$, respectively. If $X$ and $Y$ are independent then $f(x,y)_{xy} = f_x(x)f_y(y)$ and $I(X;Y) = 0$, so knowledge of one variable provides no information about the other. If $X$ and $Y$ are dependent then $I(X;Y) > 0$ and represents the expected difference in entropy between $Y$ and $Y|X$ and as such is also the expected Kullback-Leibler divergence between the marginal density $f_y$ and the conditional density $f_{y|x}$. \@ref(eq:binom)


 \begin{equation}\label{mi}
 I(X;Y) =\int \int \log \left \{ \frac{f_{xy}(x,y)}{f_{x}(x)f_{y}(y)} \right \} f_{xy}(x,y) \, dy dx, 
 \end{equation}
 where $f_x$ and $f_y$ are the marginal densities of $X$ and $Y$, respectively. If $X$ and $Y$ are independent, then $f_{xy}(x,y) = f_x(x)f_y(y)$ and $I(X;Y) = 0$, so knowledge of one variable provides no information about the other. 

For linear models with normal errors, the mutual information criterion is the same as D-optimality see appendix for the derivation. 
Maximum entropy sampling selects design points whose predictive distribution has the highest observed entropy. Suppose we have the following model $$y_i = f(x_i;\theta) + \epsilon_i$$ where $epsilon_i$ are independent and identically distributed. Then if the function $f$ is known the conditional entropy $H(Y(x)|\theta) = H(\epsilon)$ which is constant for all design points.

For sequential designs, instead of looking at mutual information between $\theta$ and $\textbf{y}$, we need to look at the mutual information between a single observation $y_n$ and $\theta$ conditioned on $\textbf{y}_{n-1}$,

\begin{eqnarray*}
I(\pi (y_n,\theta |\textbf{y}_{n-1})) &=& \int \int \int \log \left \{ \frac{\pi (y_n,\theta |\textbf{y}_{n-1})}{\pi (y_n |\textbf{y}_{n-1})\pi (\theta |\textbf{y}_{n-1})} \right \}\pi (y_n,\theta |\textbf{y}_{n-1})\pi(\textbf{y}_{n-1}) dy_n d\theta d\textbf{y}_{n-1}\\
 &=& \frac{1}{2}log \left \{ \frac{\det(\mathbf{X}_n'\mathbf{X_n} + R)}{\det(\mathbf{X}_{n-1}'\mathbf{X}_{n-1} + R)} \right \} \\
 &=& \frac{1}{2}log \left \{(1 + x'(\mathbf{X}_{n-1}'\mathbf{X}_{n-1} + R)^{-1}x) \right \},
 \end{eqnarray*}
where the last equality is derived from the matrix determinant lemma. The derivation of these results are in Appendix \ref{con-mutinfo}. Maximizing the conditional mutual information is then equivalent to maximizing

Beginning with the simplest statistical model, let us consider a linear regression model with independent and identically distributed Gaussian error with known variance $\sigma^2$. Let $\eta = \{x_i\}_{i=1}^n$ be an experimental design. The design matrix corresponding to $\eta$ will be denoted $X(\eta)$. The response is modeled as 
$$\mathbf{y} \sim norm(\mathbf{X}(\eta)\beta,\sigma^2 I_n),$$ 
where $I_n$ is the identity matrix of dimension $n$. Assuming $\beta$ has a prior distribution of $norm(0,\sigma^2 R^{-1})$ where $R$ is a positive definite matrix, then the posterior distribution of $\beta$ is
$$\beta|\mathbf{y} \sim norm((\mathbf{X}'\mathbf{X} + R)^{-1}X'\mathbf{y},\sigma^2(\mathbf{X}'\mathbf{X} + R)^{-1}).$$
The derivation of the posterior distribution is presented in Appendix~\ref{posterior_beta}. The mutual information between $\theta$ and a sample of size $n$ is
$$I( \pi(\theta,\mathbf{y})) = \int \int log \left ( \frac{\pi(\theta,\mathbf{y})}{\pi(\theta)\pi(\mathbf{y})} \right) \pi(\theta,\mathbf{y}) d\theta d\mathbf{y} = -k \log(\sigma) + \frac{1}{2}logdet(\mathbf{X}'\mathbf{X} + R).$$
The derivation of this result is presented in Appendix~\ref{mutinfo-n}. Setting $R$ to the zero matrix assumes an improper prior and shows that the D-optimal design is also the design which maximizes mutual information. This correspondence is because the model is linear in the parameter $\beta$ and the design criterion is independent of the true value of $\beta$.  

## kNN Algorithms

$$\hat{H}(X,Y) =  \psi(k) + \psi (N) + \log(c_{d_X}c_{d_Y}) + \frac{d_X+ d_Y}{N} \sum_{i=1}^N \log \epsilon(i)$$

### $KSG_1$

### $KSG_2$

### $LNC$

### $iLNC$

### Alpha Calibration
To minimize MSE we are using a GP optimization method using EQI as an acquisition function. This provides a good balance for the stochastic optimization. Other methods that were tries were: particle swarm optimization, Nelder-Mead, and Brent's method. The last two were from R's \texttt{optim} function and the swarm was from package \texttt{pso}. These methods are meant to be deterministic function optimizers so I increase the Monte Carlo sample size, but they still under performed GP optimization. 

Complicating factors include that range of the optimization. This can lead to suboptimal selection or failure to converge. We do not have any gradient information and finite-difference approximations were not found to be stable to get consistent convergence. Because the optimization is tricky, in the future we will create a surrogate to save computation time and effort that will interpolate the optimal $\alpha$ for cases we have not run through the optimizer. Depending on the values of $d$ and $k$ we will have to have to adjust the search window for the minimum.

### $LNN$ 

## Bivariate Normals

The mutual information between two 

## Functional Relationships

That paper did some functional relationship and whatnot, I can 

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Experimental Design

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
